---
title: "Bonanza Snowshoe Hares"
author: "Kiera Matiska"
date: "11/5/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach Packages
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
```

# Juvinite Hare Initial Investigation

### Read in data

- Use `read_csv()` and the here package to read in bonanza_hares.csv

```{r}
hares <- read_csv(here("data", "bonanza_hares.csv"))
```

## Annual Juvinile Hare Trap Counts

### Data Wrangling

- Create a new column, year, using the `year()` function in lubridate
- Create a new data frame using the `summarize()` function that has two columns, year and number of hare traps
- Use `kable_styling()` to finalize the table

```{r}
hares_year <- hares %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date))

hares_traps <- hares_year %>% 
  group_by(year) %>% 
  summarize(number = n())

hares_traps %>% 
  kbl() %>% 
  kable_styling()
```

### Visualize Data

Create a column graph using `ggplot()`

```{r}
ggplot(data = hares_traps, 
       aes(x = year, y = number)) +
  geom_col()
```

